Behavior Driven Development in Haskell

### 2012-05-23 paper.tex
writing solution.

\category{TBD:CR-number}{subcategory}{third-level}

\terms
TBD

\keywords
software design, documentation, automatic test

\section{Introduction}

@@

\section{Terminology}

@@

\section{Observation}

@@

\subsection{Example annotation}

@@

\subsection{Property annotation}

@@

We can see the equation as a boolean property for {\tt null} and {\tt empty} and
can say that there are potentially demands to show properties with Haddock annotation.
But there is no annotation syntax to describe properties in Haddock.
Even if property annotation syntax is defined, 
we have no agreement on how to describe non-boolean properties: e.g. how to specify types of funcation parameters.

Equation property is tested 100 times.

Quickcheck property should be able to be specified.

To our experience, it is not difficult to find properties for pure functions.
We believe that Haskellers should show properties of pure functions
in documentation.

\subsection{Internal documentation}

not useful for users

\section{Solution}

FIRST convey the idea.

\subsection{Tests suitable with doctest}

\subsection{Tests suitable with hspec}

\subsection{Integrating multiple test suites}

\section{Best current practice}

@@

\section{Implementation}

\subsection{doctest}

@@

\subsection{doctest and QuickCheck}

\subsection{hspec}

\section{Related work}

@@

\section{Conclusion}

@@

\acks

>>>>

\category{D.2.2}{Software Engineering}{Testing tools}

\terms Languages, Design, Verification

\keywords Haddock, QuickCheck, doctest, hspec

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Terminology}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Observation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Example annotation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Property annotation}

@@

We can see the equation as a boolean property for {\tt null} and {\tt empty} and
can say that there are potentially demands to show properties with Haddock annotation.
To our experience, it is not difficult to find properties for pure functions.
We believe that Haskellers should show properties of pure functions
in documentation.

But there is no annotation syntax to describe properties in Haddock.
Even if a prefix keyword for properties is defined, 
we have no agreement on how to describe non-boolean properties: e.g. how to specify types of funcation parameters. And we don't have automatic test mechanism to use properties in documentation.

Note that QuickCheck 2.3 or earlier tests a boolean property 100 times though just one test is enough. (should got to appendix?)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Internal documentation}

There are some examples and properties which are not useful for users but useful for developers. One example is a corner case registered in a bug track system. We don't have a mechanism to integrate design, documentation and tests.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Solution}

To specify visible behavior, we propose to use {\tt doctest} integrated with QuickCheck. To describe internal behavior, we propose to use {\tt hspec} aslo integrated with QuickCheck.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Visible behavior}

talk about doctest.

talk about the {\tt prop>} syntax.

talk about how to specify non-boolean properties.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Internal behavior}

TBD: talk about hspec.

\begin{quote}
\begin{verbatim}
TBD: Example
\end{verbatim}
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integrating multiple test suites}

TBD: talk about Cabal.

\begin{quote}
\begin{verbatim}
TBD: Example
Test-Suite spec
  Main-Is: Spec.hs
  Type: exitcode-stdio-1.0
  ...

Test-Suite doctests
  Main-Is: doctests.hs
  Type: exitcode-stdio-1.0
  ...
\end{verbatim}
\end{quote}

\begin{quote}
\begin{verbatim}
TBD: Spec.hs comes here
\end{verbatim}
\end{quote}

\begin{quote}
\begin{verbatim}
TBD: doctest.hs comes here
\end{verbatim}
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Best current practice}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{doctest}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{doctest and QuickCheck}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{hspec}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related work}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acks
